---
author: "Anish Bhurtyal, Avanish Chaulagai, Nikita Gerzhgorin, Joseph Donohoe"
date: "2022-11-21"
output:
  pdf_document: default
  html_document: default
  always_allow_html: yes
---


# FIFA 22 PLAYERS ANALYSIS



```{r packages, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(gghighlight)
library(knitr)
library(dplyr)
library(cluster) # clustering algorithms
library(factoextra) # clustering algorithms & visualization

set.seed(1234)
```

#### Data and exploration


### Data

```{r}
fifa <- read.csv("players_22.csv")
#head(fifa, 3)
```


### Missing values

```{r message=FALSE, warning=FALSE}
missing.values <- fifa %>%
    gather(key = "key", value = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing == T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing))

missing.values %>%
    ggplot() +
    geom_bar(aes(x = key, y = num.missing), stat = "identity") +
    labs(x = "variable", y = "number of missing values", title = "Number of missing values") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### Duplicates

```{r}
sprintf("Number of duplicates columns:")
sum(duplicated(fifa$long_name) == TRUE)
sprintf("Dimension  before")
dim(fifa)

fifa <- fifa %>% distinct(long_name, .keep_all = TRUE)
sprintf("Dimension  after")
dim(fifa)
```

### Data Cleaning

```{r datacleaning}
fifa <- fifa %>%
    mutate(wage_eur = ifelse(is.na(wage_eur),
        median(wage_eur, na.rm = T),
        wage_eur
    )) %>%
    mutate(value_eur = ifelse(is.na(value_eur),
        median(value_eur, na.rm = T),
        value_eur
    )) %>%
    mutate(passing = ifelse(is.na(passing),
        median(passing, na.rm = T),
        passing
    )) %>%
    mutate(dribbling = ifelse(is.na(dribbling),
        median(dribbling, na.rm = T),
        dribbling
    )) %>%
    mutate(pace = ifelse(is.na(pace),
        median(pace, na.rm = T),
        pace
    )) %>%
    mutate(defending = ifelse(is.na(defending),
        median(defending, na.rm = T),
        defending
    )) %>%
    mutate(shooting = ifelse(is.na(shooting),
        median(shooting, na.rm = T),
        shooting
    )) %>%
    mutate(physic = ifelse(is.na(physic),
        median(physic, na.rm = T),
        physic
    ))
```

### Visaulizing Amounts


```{r}
league1 <- table(fifa$league_name)
league2=as.data.frame(league1)
league3 <- league2 %>% 
  mutate(Freq = as.numeric(Freq)) %>% 
  arrange(desc(Freq)) %>%
  slice(1:5)
par(las=2)
par(mar=c(5,10,4,2))

barplot(league3$Freq,names.arg=c("USA Major League Soccer", "Argentina Primera División",
                                 "English League Championship", "English Premier League", "Spain Primera Division"), cex.names=0.8,xlab="Number of Players",col="blue",
main="Top 5 Leagues with Highest Number of Players",border="red", horiz=TRUE)
```

### Where are the players from?

```{r}
fifa_country_count <- fifa %>%
    group_by(nationality_name) %>%
    summarize(Freq = n())

fifa_country_count$nationality_name[fifa_country_count$nationality_name 
                                    == "United States"] <- "United States of America"
fifa_country_count
```

```{r message=FALSE, warning=FALSE, fig.width=9, fig.height= 7} 
library(plotly)
country_with_code <- read.csv("https://raw.githubusercontent.com/lukes/ISO-3166-Countries-with-Regional-Codes/master/slim-3/slim-3.csv")
head(country_with_code)

fifa_country_count_with_code <- fifa_country_count %>%
    left_join(country_with_code,
        by = c("nationality_name" = "name")
    )

fig <- plot_ly(fifa_country_count_with_code,
    type = "choropleth",
    locations = fifa_country_count_with_code$alpha.3,
    z = fifa_country_count_with_code$Freq,
    text = fifa_country_count_with_code$nationality_name,
    colorscale = "ice"
)

fig <- fig %>% colorbar(title = "No of players")
fig <- fig %>% layout(
    title = "Choropleth showing FIFA 2022 Players' Nationality <br>(Hover for breakdown)"
)

fig
```





#### Visaulizing Proportions

```{r}
nation1 <- table(fifa$nationality_name)
nation2=as.data.frame(nation1)
nation3 <- nation2 %>% 
  mutate(Freq = as.numeric(Freq)) %>% 
  arrange(desc(Freq)) %>%
  slice(1:5)
pct <- round(nation3$Freq/sum(nation3$Freq)*100)
nation3$Var1 <- paste(pct,"%",sep="") 
colors <- c("grey", "blue","green", "yellow", "red")
pie(nation3$Freq,labels = nation3$Var1, col = colors,
    main="Top 5 Countries with Highest Number of Players", cex=0.8)
legend(1.2, .5, c("England", "Germany", "Spain", "France", "Argentina"), fill = colors) 
```





### Age Distribution


```{r warning=FALSE, fig.asp = 0.25, fig.width= 7}
ggplot(fifa, aes(x = age)) +
    geom_boxplot() +
    labs(
        title = "The Distribution of Age",
        x = "Age"
    )
```


### Oldest Players


```{r}
old <- fifa %>%
    arrange(desc(age)) %>%
    select(short_name, nationality_name, age)
head(old)
```


### Youngest Players


```{r}
young <- fifa %>%
    arrange(age) %>%
    select(short_name, nationality_name, age)
head(young)
```

### Some line charts with age

### Players Attributes vs. Age

```{r message=FALSE, warning=FALSE}
filtered_attributes <- fifa %>%
    group_by(age) %>%
    summarise_at(
        vars(overall, potential, pace),
        list(mean)
    )
filtered_attributes

library("reshape2")
data_long <- melt(filtered_attributes, id = "age")

ggplot(
    data_long,
    aes(
        x = age,
        y = value,
        color = variable
    )
) +
    geom_line() +
    labs(
        y = "Player Atrributes Rating",
        x = "Age",
        color = "Color Legend",
        title = "Attributes vs Age",
    )
```


### Players Value vs. Age

```{r}

money_attributes <- fifa %>%
    group_by(age) %>%
    summarise_at(
        vars(value_eur),
        list(mean)
    )

ggplot(money_attributes, aes(x = age, y = value_eur)) +
    geom_line() +
    labs(
        y = "Player Value in Millions",
        x = "Age",
        title = "Player value vs Age",
    ) +
    scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) # thousands
```



#### Modeling

```{r}
hist(fifa$overall, xlab="overall Rating", main = "Histogram of Overall Rating")
fifa$overall = log10(fifa$overall)


hist(fifa$overall, xlab = "Log10 of Overall Rating", main = "Histogram of Overall Rating- Log10")

hist(fifa$wage_eur, xlab = "Wage in Euros", main = "Histogram of Wage")
fifa$wage_eur = log10(fifa$wage_eur)

#fifa$wage_eur = log(fifa$overall)
hist(fifa$overall, xlab="Wage in Euros", main="Histogram of Wage- Log10")

```

### Train / test

- Create an initial split

```{r}
dt <- sort(sample(nrow(fifa), nrow(fifa) * .7))
train <- fifa[dt, ]
test <- fifa[-dt, ]

train <- train %>% filter(!is.na(overall))
test <- test %>% filter(!is.na(overall))

```


```{r}


overall_individualskills <- lm(overall ~ pace + shooting + passing + dribbling + defending +
                                 physic + age + preferred_foot + attacking_crossing + attacking_finishing + attacking_heading_accuracy + 
                                 attacking_short_passing + attacking_volleys + skill_dribbling +skill_curve +skill_fk_accuracy +
                                 skill_long_passing +skill_ball_control +movement_acceleration +movement_sprint_speed +
                                 movement_agility +movement_reactions +movement_balance +power_shot_power +power_jumping+ power_stamina +
                                 power_strength +power_long_shots +mentality_aggression +mentality_interceptions +mentality_positioning +
                                 mentality_vision +mentality_penalties +mentality_composure +defending_marking_awareness +defending_standing_tackle +
                                 defending_sliding_tackle, data = train)

summary(overall_individualskills)



wage_individualskills <- lm(wage_eur ~ pace + shooting + passing + dribbling + defending +
                              physic + age + preferred_foot + attacking_crossing + attacking_finishing + attacking_heading_accuracy +
                              attacking_short_passing + attacking_volleys + skill_dribbling +skill_curve +skill_fk_accuracy +
                              skill_long_passing +skill_ball_control +movement_acceleration +movement_sprint_speed +movement_agility +
                              movement_reactions +movement_balance +power_shot_power +power_jumping+ power_stamina +
                              power_strength +power_long_shots +mentality_aggression +mentality_interceptions +mentality_positioning +
                              mentality_vision +mentality_penalties +mentality_composure +defending_marking_awareness +defending_standing_tackle +
                              defending_sliding_tackle, data = train)

summary(wage_individualskills)

```



```{r}

overalldata <- data.frame(actual= test$overall, predicted = predict(overall_individualskills, test))
wagedata <- data.frame(actual= test$wage_eur, predicted = predict(wage_individualskills, test))
#colSums(is.na(data2))
```

```{r}
m2wage <- mean((wagedata$actual- wagedata$predicted)^2)
m2wage
m2overall <-  mean((overalldata$actual - overalldata$predicted)^2)
m2overall
```
```{r  out.width = "70%"}
ggplot(wagedata, aes(x = predicted, y = actual)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  labs(x='Transformed Predicted Wage Values', y='Transformed Actual Wage Values',
       title='Transformed Predicted vs. Transformed Actual Wage Values')

ggplot(overalldata, aes(x = predicted, y = actual)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  labs(x='Transformed Predicted Overall Values', y='Transformed Actual Overall Values',
       title='Transformed Predicted vs. Transformed Actual Overall Values')

```



### Ghana players cluster


```{r fig.width=9, fig.height= 7}
# filter out players from team Ghana who arenot reserves or substitutes
ghana <- fifa %>%
    filter(fifa$nationality_name == "Ghana" & fifa$club_position != "RES" &
             fifa$club_position != "SUB") %>%
    select(short_name, shooting, pace, passing, dribbling, defending)

# drop duplicates if any
ghana <- ghana %>% distinct(short_name, .keep_all = TRUE)

# make player name as rownames
rownames(ghana) <- ghana$short_name
ghana2 <- ghana %>% select(2:5)

# perform k means clustering with 10 centers
k2 <- kmeans(ghana2, centers = 10, nstart = 25)
# str(k2)

fviz_cluster(k2, data = ghana2)
```


### Current Portugal Team Cluster


```{r  fig.width=9, fig.height= 7 }
portugal_current <- c(
    "André Silva", "Antonio Silva", "Bernardo Silva", "Bruno Fernandes",
    "Cristiano Ronaldo", "Danilo Pereira", "Diogo Costa", "Diogo Dalot",
    "Gonçalo Ramos", "João Cancelo", "João Félix", "João Mário", "Palhinha",
    "José Sá", "Matheus Nunes", "Nuno Mendes", "Otavio", "Pepe", "Rafael Leão",
    "Raphael Guzzo", "Ricardo Horta", "Rúben Dias", "Ruben Neves", "Rui Patrício",
    "Vitinha", "William Carvalho"
)

portugal <- fifa %>%
    filter(fifa$nationality_name == "Portugal") %>%
    select(short_name, shooting, pace, passing, dribbling, defending)

# filter out the players in current national team
portugal <- portugal[portugal$short_name %in% portugal_current, ]

# drop duplicates if any
portugal <- portugal %>% distinct(short_name, .keep_all = TRUE)

# make player name as rownames
rownames(portugal) <- portugal$short_name
portugal2 <- portugal %>% select(2:5)

# perform k means clustering with 5 centers
k2 <- kmeans(portugal2, centers = 5, nstart = 25)
# str(k2)
fviz_cluster(k2, data = portugal2)
```

```{r warning=FALSE, fig.width=9, fig.height= 7 }
res <- hcut(portugal2, k = 4, stand = TRUE)
# Visualize
fviz_dend(res, rect = TRUE, cex = 0.5,
          k_colors = c("#00AFBB","#2E9FDF", "#E7B800", "#FC4E07"))
```

